The Hyperedge Event Model

### 2018-05-20 paper/ba-HEM/ba-sample.tex
PPE results updated

	\begin{algorithm}[ht]

@@

New values for continuous variables $\boldsymbol{b},$ and $\boldsymbol{\eta}$ and $\sigma^2_\tau$ (if applicable) cannot be sampled directly from their conditional posteriors, but may instead be obtained using the Metropolis--Hastings algorithm. With uninformative priors (i.e., $N({0},\infty)$), the conditional posterior over $\boldsymbol{b}$ is
\begin{equation}
P(\boldsymbol{b}| \boldsymbol{u}, \boldsymbol{x})\propto \prod_{d=1}^D
\prod_{a=1}^A \frac{\exp\Big\{\mbox{log}\big(\text{I}( \lVert \boldsymbol{u}_{ad}\rVert_1 > 0)\big) + \sum\limits_{r \neq a} \lambda_{adr}u_{adr}\Big\}}{Z(\boldsymbol{\lambda}_{ad})},
\end{equation}
where the two variables share the conditional posterior and thus can be jointly sampled. Likewise, assuming uninformative priors on $\boldsymbol{\eta}$ (i.e., $N({0},\infty)$) and $\sigma_{\tau}^2$ (i.e., half-Cauchy($\infty$)), the conditional posterior for no-tied event case is
\begin{equation}
P(\boldsymbol{\eta}, \sigma_\tau^2| \boldsymbol{u}, \boldsymbol{y})\propto \prod_{d=1}^D\Big(f_{\tau}(\tau_{d}; \mu_{a_d d}, \sigma_\tau^2)\times \prod_{a\neq a_d}\big(1-F_{\tau}(\tau_{d}; \mu_{a d}, \sigma_\tau^2) \big)\Big),
\end{equation}
where $f_{\tau}(\tau_{d}; \mu_{a_d d}, \sigma_\tau^2)$ is the probability that the $d^{th}$ observed time-increment comes from the specified distribution $f_\tau(\cdot)$ with the observed sender's mean parameter, and $\prod_{a\neq a_d}\big(1-F_{\tau}(\tau_{d}; \mu_{a d}, \sigma_\tau^2) \big)$ is the probability that the latent (or unobserved) senders for event $d$ all generated time-increments greater than $\tau_d$. Moreover, under the existence of tied-event, the conditional posterior of $\boldsymbol{\eta}$ and $\sigma_{\tau}^2$ should be written as
\begin{equation}
\begin{aligned}
&P(\boldsymbol{\eta}, \sigma_\tau^2| \boldsymbol{u}, \boldsymbol{y})\\&\propto \prod_{m=1}^M\Big(\prod_{d:t_d=t_m^*}f_{\tau}(t_m^*-t_{m-1}^*; \mu_{a_d d}, \sigma_\tau^2) \times \prod_{a \notin \{a_d\}_{d:t_d=t_m^*}}\big(1-F_{\tau}(t_m^*-t_{m-1}^*; \mu_{a d}, \sigma_\tau^2) \big)\Big),
\end{aligned}
\end{equation}
where $t_1^*,\ldots,t_M^*$ are the unique timepoints across $D$ events ($M \leq D$). If $M=D$ (i.e., no tied events), equation (3.4) reduces to equation (3.3). Algorithm \ref{alg:MCMC} provides the pseudocode.
\begin{algorithm}[H]
	\spacingset{1}
	\SetAlgoLined
	\caption{MCMC Algorithm}
	\begin{algorithmic}
	\STATE set initial values of $\boldsymbol{b}, \boldsymbol{\eta}$, (and $\sigma_\tau^2$)\\
	\FOR{o=1 to outer}
		\FOR{d=1 to D}
		 \FOR{a = 1 to A}
			\FOR{r = 1 to A (r $\neq$ a)}
					\STATE update $u_{adr}$ using Gibbs update --- equation (7)
			\ENDFOR
		\ENDFOR
		\ENDFOR
		\FOR{n=1 to inner1}
			\STATE update $\boldsymbol{b}$ using Metropolis-Hastings --- equation (8)
		\ENDFOR
	\FOR{n=1 to inner2}
			\STATE update $\boldsymbol{\eta}$ using Metropolis-Hastings --- equation (10)
			\STATE (if needed) update $\sigma_\tau^2$ using Metropolis-Hastings --- equation (10)
		\ENDFOR

@@

   In the forward samples, both observable and unobservable variables are generated using Algorithm \ref{alg:generative}. In the backward samples, unobservable variables are generated using the sampling equations for inference, which we derived in Section \ref{subsec:conditionaldist}. For each forward and backward sample that consists of $D$ number of edges, we save these statistics:
   \begin{itemize}
   	\item[1.] Mean of observed recipient sizes $ \lVert \boldsymbol{r}_{d} \rVert_1 $ across $d=1,\ldots,D$,
   	\item[2.] Mean of time-increments $\tau_d$ across $d=1,...,D$,
   	\item[3.] $b_p$ value used to generate the samples $p = 1,...,P$,
   	\item[4.] $\eta_p$ value used to generate the samples $p = 1,...,P$,
   	\item[5.] $\sigma^2_\tau$ value used to generate the samples
   \end{itemize}

   To keep the computational burden of re-running thousands of rounds of inference manageable, we run the GiR using a relatively small artificial sample, consisting of $D=100$ edges, $A=5$ actors, $P=4$ number of edge covariates, and $Q=3$ number of timestamp covariates per each forward or backward samples, using lognormal distibution for the time distribution $f$. We generated $10^5$ sets of forward and backward samples, and then calculated 1,000 quantiles for each of the statistics. We also calculated t-test and Mann-Whitney test p-values in order to test for differences in the distributions generated in the forward and backward samples. Before we calculated these statistics, we thinned our samples 
   by taking every 9th sample starting at the 10,000th sample for a resulting sample size of 10,000, in order to reduce the autocorrelation in the Markov chains. In each case, if we observe a large p-value, this gives us evidence that the distributions generated under forward and backward sampling have the same locations. We depict the GiR results using probability-probability (PP) plots. To compare two samples with a PP-plot we calculate the empirical quantile in each sample of a set of values observed across the two samples, then plot the sets of quantiles in the two samples against each other. If the two samples are from equivalent distributions, the quantiles should line up on a line with zero $y$-intercept, and unit slope (i.e., a 45-degree line). The GiR test results are depicted in Figure \ref{fig:PPplot}, which show that we pass the test on every statistic.
   \begin{figure}[ht]
   	\centering
   			\includegraphics[width=1\textwidth]{img/etaest-1.png}	
   	\label{fig:PPplot}
   	\caption{Probability-Probability plot for the GiR test statistics.}
   \end{figure}

@@

 	   We now present a case study applying our method to Montgomery county government email data. For this case study,

@@

 	   \subsection{Data}\label{subsec:Montgomery}
 	   Our data come from the North Carolina county government email dataset collected by \cite{ben2017transparency} that includes internal email corpora covering the inboxes and outboxes of managerial-level employees of North Carolina county governments. Out of over twenty counties, we chose Montgomery County to 1) test our model using data with large portion of emails with hyperedges (16.76\%), all of which are a single sender multiple receiver cases, and 2) limit the scope of this initial application. To summarize, Montgomery County email network contains 680 emails, sent and received by 18 department managers over a period of 3 months (March--May) in 2012. 

@@

	We evaluate the model's ability to predict edges and timestamps from the Montgomery County email data, conditioned on their ``training" part of the data. To perform the experiment, we separately formed a test split of each three components---sender, recipients, and timestamps---by randomly selecting ``test" data with probability $p=0.1$. Any missing variables were imputed by drawing samples from their conditional posterior distributions, given the observed data, parameter estimates, and current values of test data. We then run inference to update the latent variables given the imputed and observed data. We iterate the two steps---imputation and inference---multiple times to obtain enough number of estimates for ``test" data. Algorithm \ref{alg:PPE} outlines this procedure in detail. We run the experiment and measure the predictive performance of two separate time distributions using 500 predicted samples, which are depicted in Figure \ref{figure:PPEresults}. First, we compare the correct sender probability for missing documents $\{d:a_d=\mbox{NA}\}$, which corresponds to $\pi_{a_{d}}=P(a_{d} = a^{obs}_{d}|\cdot)$ in Algorithm \ref{alg:PPE}. The plot on the left shows that lognormal distribution reaches higher correct sender probability than exponential distribution, although their centers (e.g., medians) are similar. Secondly, we compute similar measure for recipients---correct recipient probability $P(r_{dr} = r^{obs}_{dr})$ in Algorithm \ref{alg:PPE} where $r^{obs}_{dr}$ is either 1 (received) or 0 (not received)---and visually compare the two distributions in logit scale (i.e., $\log(\frac{P(r_{dr} = r^{obs}_{dr}|\cdot)}{1-P(r_{dr} = r^{obs}_{dr})})$). Since the edge generating process (Section \ref{subsec: Tie}) is not directly affected by the choice of timestamp distribution, there is no significant difference between lognormal and exponential in their performance of predicting unknown recipients. Finally, prediction errors for missing timestamps are measured using the median absolute percentage error (MdAPE) \citep{hyndman2006another}, where we take the median of absolute percentage errors $p_{di}$ for $i=1,\ldots,500$ for $d$th document, calculated as follows:
	\begin{figure}[!t]
		\centering
		\includegraphics[width=1\textwidth]{img/PPEplot-1.png}	
		\caption {Comparison of predictive performance between lognormal and exponential distributions: boxplots for correct sender probability (\textit{left}), distribution of correct recipient probability in logit (\textit{middle}), and boxplots for median absolute percentage error (\textit{right}).}
		\label{figure:PPEresults}
	\end{figure}
	\begin{equation*}
	p_{di} = \abs*{\frac{\tau^{obs}_d - \tau^{pred}_{di}}{\tau^{obs}_d}},
	\end{equation*}
	and we present boxplots for the MdAPE estimates on the right plot. As expected, we have huge benefit in the performance of timestamp prediction when we choose lognormal distribution over exponential. This difference can be explained by our variance estimate $\hat{\sigma}_\tau^2= 14.093$ in Section \ref{subsec:Result_email}, because exponential distribution omits the extra information on time-increments contained in the large variance estimate. As illustrated above, we can use this out-of-sample prediction task for two usage---1) provide an effective answer to the question ``how are we filling in the missing information of the emails?" and 2) offer one standard way to determine the time-increment distribution in Section \ref{subsec:Time}.

@@

	Figure \ref{figure:PPCresults} illustrates the results of poterior predictive checks using lognormal distribution. Upper two plots show node-specific degree distributions, where the left one is the boxplots of 500 posterior predictive outdegree statistic and the right plot is the same one for indegree statistic. For both plots, the x-axis represents the node id's from 1 to 18, and the y-axis represents the number of emails sent or received by the node. When compared with the observed outdegree and indegree statistics (red lines), our model recovers the overall distribution of sending and receiving activities across the nodes. For example, node 1 and 10 show significantly higher level of both sending and receiving activities relative to the rest, and the model-simulated data captures those big jumps, without using node-specific indicators. Outdegree distribution of some low-activity nodes are not precisely recovered, however, indegree distribution looks much better. Since we use more information in the recipient selection process (i.e., network effects) while we only rely on minimum time-increments when choosing the observed sender, these results are expected. Lower left plot is the distribution of recipient sizes, where the x-axis spans over the size of recipients 1 to 14 (which is the maximum size of observed recipients) and the y-axis denotes the number of documents with x-number of recipients. The result shows that our model is underestimating single-receiver documents while overestimating documents with two, three, and four recipients. One explanation behind what we observe is that the model is trying to recover so-called ``broadcast'' emails, which are the documents with $\geq 10$ number of recipients, so that the intercept estimate $b_1$ is slightly moved toward right. To improve the goodness of fit for recipient sizes, we can add more covariates (e.g., sender indicators) or assign strong prior structure on $b_1$. In the end, the plot on the lower right is the PP plot for time-increments, which is a graphical measure commonly used for assessing how closely two data sets agree, by plotting the two cumulative distribution functions against each other. The resulting goodness of fit of the diagonal line gives a measure of the difference between the distribution of simulated time-increments and the observed ones, and our PP plot shows that we have great performance in reproducing the observed time distribution. All our findings from predictive experiments in Section \ref{subsec:Experiment_email} are further revealed in the PPC results from exponential distribution, which are attached in Appendix B.
	\subsection{Parameter Estimates and Interpretation}\label{subsec:Result_email}
	Based on some preliminary experiments, we choose lognormal distribution to model time-to-next-email in Montgomery county email data, thus we estimate the variance parameter $\sigma_\tau^2$ with the prior specified as inverse-Gamma($2, 1$). In this section, we present the results based on 55,000 MCMC outer iterations with a burn-in of 15,000, where we thin by keeping every 40th sample. While the inner iterations for $\boldsymbol{\eta}$ and $\sigma_\tau^2$ are fixed as 1, we use 10 inner iterations for $\boldsymbol{b}$ to adjust for its slower convergence rate.
	\begin{figure}[!b]
		\centering
		\includegraphics[width=1\textwidth]{img/best-1.png}	
		\caption {Posterior distribution of $\boldsymbol{b}$ estimates.}
		\label{figure:betaresults}
	\end{figure}

@@

	To use this distribution efficiently, we derive a closed-form expression for $Z(\boldsymbol{\lambda}_{id})$ that does not require brute-force summation over the support of $\boldsymbol{u}_{ad}$ (\textit{i.e.} $\forall \boldsymbol{u}_{ad} \in [0,1]^A$). We recognize that if $\boldsymbol{u}_{ad}$ were drawn via independent Bernoulli distributions in which $P({u}_{adr}=1|\delta, \boldsymbol{\lambda}_{ad})$ was given by logit$(\lambda_{adr})$, then 
	\begin{equation*}
	P(\boldsymbol{u}_{ad}|\boldsymbol{\lambda}_{ad}) \propto \exp\Big\{\sum_{r \neq a } \lambda_{adr}u_{adr}\Big\}.  	 
	\end{equation*}
	This is straightforward to verify by looking at 
	\begin{equation*}
	\begin{aligned}
	&P(u_{adr}=1|\boldsymbol{u}_{ad[-r]}, \boldsymbol{\lambda}_{ad})
	=\frac{ \exp{(\lambda_{adr})}}{\exp{(\lambda_{adr})} + 1}.\end{aligned}\end{equation*}
	We denote the logistic-Bernoulli normalizing constant as $Z^{l}(\boldsymbol{\lambda}_{ad})$, which is defined as 
	\begin{equation*}
	Z^{l}(\delta,\boldsymbol{\lambda}_{ad})=\sum_{\boldsymbol{u}_{ad} \in [0,1]^{A}} \exp\Big\{\sum_{r\neq a} \lambda_{adr}u_{adr}\Big\}.
	\end{equation*}
	Now, since 
	\begin{equation*}
	\begin{aligned}
	&\exp\Big\{ \mbox{log}\Big(\text{I}(\lVert \boldsymbol{u}_{ad} \rVert_1 > 0)\Big) + \sum_{r \neq a} \lambda_{adr}u_{adr} \Big\}\\&= \exp\Big\{  \sum_{r \neq a} \lambda_{adr}u_{adr} \Big\},
	\end{aligned}
	\end{equation*}
	except when $\lVert \boldsymbol{u}_{ad} \rVert_1=0$, we note that 
	\begin{equation*}
	\begin{aligned}
	Z(\boldsymbol{\lambda}_{ad})& = Z^{l}(\boldsymbol{\lambda}_{ad}) -\exp\Big\{ \sum\limits_{\forall u_{adr}=0}\lambda_{adr}u_{adr} \Big\}
	\\& = Z^{l}(\lambda_{a}^{(d)}) -  1.
	\end{aligned}
	\end{equation*}
	We can therefore derive a closed form expression for $Z(\delta,\boldsymbol{\lambda}_{ad})$ via a closed form expression for $Z^{l}(\delta,\boldsymbol{\lambda}_{ad})$. This can be done by looking at the probability of the zero vector under the logistic-Bernoulli model:
	\begin{equation*}
	\begin{aligned}
	&\frac{\exp\Big\{ \sum\limits_{\forall u_{adr}=0}\lambda_{adr}u_{adr} \Big\}}{Z^{l}(\boldsymbol{\lambda}_{ad})}= \prod_{r \neq a}   \Big(1-\frac{ \exp{(\lambda_{adr})}}{\exp{(\lambda_{adr})} + 1}\Big).
	\end{aligned}  
	\end{equation*}
	Then, we have 
	\begin{equation*}
	\begin{aligned}
	& \frac{1}{Z^{l}(\boldsymbol{\lambda}_{ad})} &= \prod\limits_{r \neq a}\frac{1}{ \exp(\lambda_{adr})+ 1}.
	\end{aligned}  
	\end{equation*}
	Finally, the closed form expression for the normalizing constant is  \begin{equation*}
	\begin{aligned}Z(\boldsymbol{\lambda}_{ad}) = \prod_{r \neq a } \big(\mbox{exp}(\lambda_{adr}) + 1\big)-1.
	\end{aligned}  
	\end{equation*}
	\subsection*{Appendix B: PPC Results from Exponential Distribution}\label{subsec: PPCexp}
	\begin{figure}[H]
		\centering
		\includegraphics[width=0.495\textwidth]{img/outdegree_exp-1.png}	
		\includegraphics[width=0.495\textwidth]{img/indegree_exp-1.png}	
		\includegraphics[width=0.495\textwidth]{img/recipientsize_exp-1.png}	
		\includegraphics[width=0.31\textwidth]{img/timepp_exp-1.png}
		\caption {PPC results from exponential distribution: outdegree distribution (\textit{upper left}), indegree distribution (\textit{upper right}), recipient size distribution (\textit{lower left}), and time-increments probability-probability (PP) plot  (\textit{lower right}). Red lines in the first three plot depict the observed statistics, and the red line in the last plot is the diagonal line connecting $(0, 0)$ and $(1, 1)$.}
		\label{figure:PPCexp}
	\end{figure}

>>>>

	\begin{algorithm}[!b]

@@

New values for continuous variables $\boldsymbol{b},$ and $\boldsymbol{\eta}$ and $\sigma^2_\tau$ (if applicable) cannot be sampled directly from their conditional posteriors, but may instead be obtained using the Metropolis--Hastings algorithm. With uninformative priors (i.e., $N({0},\infty)$), the conditional posterior over $\boldsymbol{b}$ is
\begin{equation}
P(\boldsymbol{b}| \boldsymbol{u}, \boldsymbol{x})\propto \prod_{d=1}^D
\prod_{a=1}^A \frac{\exp\Big\{\mbox{log}\big(\text{I}( \lVert \boldsymbol{u}_{ad}\rVert_1 > 0)\big) + \sum\limits_{r \neq a} \lambda_{adr}u_{adr}\Big\}}{Z(\boldsymbol{\lambda}_{ad})},
\end{equation}
where the two variables share the conditional posterior and thus can be jointly sampled. Likewise, assuming uninformative priors on $\boldsymbol{\eta}$ (i.e., $N({0},\infty)$) and $\sigma_{\tau}^2$ (i.e., half-Cauchy($\infty$)), the conditional posterior for no-tied event case is
\begin{equation}
P(\boldsymbol{\eta}, \sigma_\tau^2| \boldsymbol{u}, \boldsymbol{y})\propto \prod_{d=1}^D\Big(f_{\tau}(\tau_{d}; \mu_{a_d d}, \sigma_\tau^2)\times \prod_{a\neq a_d}\big(1-F_{\tau}(\tau_{d}; \mu_{a d}, \sigma_\tau^2) \big)\Big),
\end{equation}
where $f_{\tau}(\tau_{d}; \mu_{a_d d}, \sigma_\tau^2)$ is the probability that the $d^{th}$ observed time-increment comes from the specified distribution $f_\tau(\cdot)$ with the observed sender's mean parameter, and $\prod_{a\neq a_d}\big(1-F_{\tau}(\tau_{d}; \mu_{a d}, \sigma_\tau^2) \big)$ is the probability that the latent (or unobserved) senders for event $d$ all generated time-increments greater than $\tau_d$. Moreover, under the existence of tied-event, the conditional posterior of $\boldsymbol{\eta}$ and $\sigma_{\tau}^2$ should be written as
\begin{equation}
\begin{aligned}
&P(\boldsymbol{\eta}, \sigma_\tau^2| \boldsymbol{u}, \boldsymbol{y})\\&\propto \prod_{m=1}^M\Big(\prod_{d:t_d=t_m^*}f_{\tau}(t_m^*-t_{m-1}^*; \mu_{a_d d}, \sigma_\tau^2) \times \prod_{a \notin \{a_d\}_{d:t_d=t_m^*}}\big(1-F_{\tau}(t_m^*-t_{m-1}^*; \mu_{a d}, \sigma_\tau^2) \big)\Big),
\end{aligned}
\end{equation}
where $t_1^*,\ldots,t_M^*$ are the unique timepoints across $D$ events ($M \leq D$). If $M=D$ (i.e., no tied events), equation (3.4) reduces to equation (3.3). Algorithm \ref{alg:MCMC} provides the pseudocode.
\begin{algorithm}[!b]
	\spacingset{1}
	\SetAlgoLined
	\caption{MCMC Algorithm}
	\begin{algorithmic}
	\STATE set initial values of $\boldsymbol{b}, \boldsymbol{\eta}$, (and $\sigma_\tau^2$)\\
	\FOR{o=1 to outer}
		\FOR{d=1 to D}
		 \FOR{a = 1 to A}
			\FOR{r = 1 to A (r $\neq$ a)}
					\STATE update $u_{adr}$ using Gibbs update --- equation (7)
			\ENDFOR
		\ENDFOR
		\ENDFOR
		\FOR{n=1 to inner1}
			\STATE update $\boldsymbol{b}$ using Metropolis-Hastings --- equation (8)
		\ENDFOR
	\FOR{n=1 to inner2}
			\STATE update $\boldsymbol{\eta}$ using Metropolis-Hastings --- equation (10)
			\STATE (if needed) update $\sigma_\tau^2$ using Metropolis-Hastings --- equation (10)
		\ENDFOR

@@

   In the forward samples, both observable and unobservable variables are generated using Algorithm \ref{alg:generative}. In the backward samples, unobservable variables are generated using the sampling equations for inference, which we derived in Section \ref{subsec:conditionaldist}. For each forward and backward sample that consists of $D$ number of edges, we save these statistics:
   \begin{itemize}
   	\item[1.] Mean of observed recipient sizes $ \lVert \boldsymbol{r}_{d} \rVert_1 $ across $d=1,\ldots,D$,
   	\item[2.] Mean of time-increments $\tau_d$ across $d=1,...,D$,
   	\item[3.] $b_p$ value used to generate the samples $p = 1,...,P$,
   	\item[4.] $\eta_q$ value used to generate the samples $q = 1,...,Q$,
   	\item[5.] $\sigma^2_\tau$ value used to generate the samples
   \end{itemize}

   To keep the computational burden of re-running thousands of rounds of inference manageable, we run the GiR using a relatively small artificial sample, consisting of $D=100$ edges, $A=5$ actors, $P=4$ number of edge covariates, and $Q=3$ number of timestamp covariates per each forward or backward samples, using lognormal distibution for the time distribution $f$. We generated $10^5$ sets of forward and backward samples, and then calculated 1,000 quantiles for each of the statistics. We also calculated t-test and Mann-Whitney test p-values in order to test for differences in the distributions generated in the forward and backward samples. Before we calculated these statistics, we thinned our samples by taking every 9th sample starting at the 10,000th sample for a resulting sample size of 10,000, in order to reduce the autocorrelation in the Markov chains. In each case, if we observe a large p-value, this gives us evidence that the distributions generated under forward and backward sampling have the same locations. We depict the GiR results using probability-probability (PP) plots. To compare two samples with a PP-plot we calculate the empirical quantile in each sample of a set of values observed across the two samples, then plot the sets of quantiles in the two samples against each other. If the two samples are from equivalent distributions, the quantiles should line up on a line with zero $y$-intercept, and unit slope (i.e., a 45-degree line). The GiR test results are depicted in Figure \ref{figure:GiRplot}, which show that we pass the test on every statistic.
   \begin{figure}[ht]
   	\includegraphics[width=1\textwidth]{img/GiRplot-1.png}	
   	\caption{Probability-Probability plot for the GiR test statistics.}
   	   	\label{figure:GiRplot}

@@

 	   We now present a case study applying our method to Montgomery county government email data.
 	   %\subsection{Data}\label{subsec:Montgomery}
 	   Our data come from the North Carolina county government email dataset collected by \cite{ben2017transparency} that includes internal email corpora covering the inboxes and outboxes of managerial-level employees of North Carolina county governments. Out of over twenty counties, we chose Montgomery County to 1) test our model using data with large portion of emails with hyperedges (16.76\%), all of which are a single sender multiple receiver cases, and 2) limit the scope of this initial application. To summarize, Montgomery County email network contains 680 emails, sent and received by 18 department managers over a period of 3 months (March--May) in 2012.  For this case study,

@@

@@

	We evaluate the model's ability to predict edges and timestamps from the Montgomery County email data, conditioned on their ``training" part of the data. To perform the experiment, we separately formed a test split of each three components---sender, recipients, and timestamps---by randomly selecting ``test" data with probability $p=0.1$. Any missing variables were imputed by drawing samples from their conditional posterior distributions, given the observed data, parameter estimates, and current values of test data. We then run inference to update the latent variables given the imputed and observed data. We iterate the two steps---imputation and inference---multiple times to obtain enough number of estimates for ``test" data. Algorithm \ref{alg:PPE} outlines this procedure in detail. We run the experiment and measure the predictive performance of two separate time distributions using 500 predicted samples, which are depicted in Figure \ref{figure:PPEresults}. We compare the predictions in terms of classification accuracy in predicting the authors and recipients, as well as prediction error in the timestamps. First, we compare the posterior probability of correct senders from missing documents $\{d:a_d=\mbox{NA}\}$, which corresponds to $\pi_{a_{d}}=P(a_{d} = a^{obs}_{d}|\cdot)$ in Algorithm \ref{alg:PPE}. We call this measure as ``correct sender posterior probability." The plot on the left demonstrates that both lognormal and exponential distributions acheive better predictive performance for senders compared to what is expected under random guess (dotted blue line at $1/18$), with lognormal distribution showing slightly higher performance than exponential distribution on average. Secondly, since the recipient vector is binary---$r_{dr}=1$ if received or $r_{dr}=0$ otherwise---we compute $F_1$ scores for recipient predictions, comparing the true recipient vector $\boldsymbol{r}^{obs}_{dr}$ and predicted recipient vectors $\boldsymbol{r}^{pred}_{dr}$ for missing recipient indicators (i.e., all $d$ and $r$ with $r_{dr}$=NA). Although the edge generating process (Section \ref{subsec: Tie}) is not directly affected by the choice of timestamp distribution, there exists noticeable difference between lognormal and exponential in their performance of predicting unknown recipients. Similar to the sender prediction, lognormal outperforms exponential in general. Finally, prediction errors for missing timestamps are measured using the median absolute percentage error (MdAPE) \citep{hyndman2006another}, where we take the median of absolute percentage errors $p_{di}$ for $i=1,\ldots,500$ for $d$th document, calculated as follows:
		\includegraphics[width=1\textwidth]{img/PPEplotnew-1.png}	
		\caption {Comparison of predictive performance between lognormal and exponential distributions: boxplots for correct sender posterior probability (\textit{left}), distribution of correct recipient posterior probability in log-scale (\textit{middle}), and boxplots for median absolute percentage error (\textit{right}). The dotted blue lines in the left plot represents the correct sender probability expected by random guess---i.e., $1/A=1/18\approx0.056$.}
		\label{figure:PPEresults}
	\begin{equation*}
	p_{di} = \abs*{\frac{\tau^{obs}_d - \tau^{pred}_{di}}{\tau^{obs}_d}},
	\end{equation*}
	and we present boxplots for the MdAPE estimates on the right plot. For better visualization, we plot MdAPE estimates in log-scale since they are highly right-skewed. As expected, we have huge benefit in the performance of timestamp prediction when we choose lognormal distribution over exponential. This difference can be explained by our variance estimate $\hat{\sigma}_\tau^2= 14.093$ in Section \ref{subsec:Result_email}, because exponential distribution omits the extra information on time-increments contained in the large variance estimate. As illustrated above, we can use this out-of-sample prediction task for two usage---1) provide an effective answer to the question ``how are we filling in the missing information of the emails?" and 2) offer one standard way to determine the time-increment distribution in Section \ref{subsec:Time}.

@@

	Figure \ref{figure:PPCresults} illustrates the results of poterior predictive checks using lognormal distribution. Upper two plots show node-specific degree distributions, where the left one is the boxplots of 500 posterior predictive outdegree statistic and the right plot is the same one for indegree statistic. For both plots, the x-axis represents the node id's from 1 to 18, and the y-axis represents the number of emails sent or received by the node. When compared with the observed outdegree and indegree statistics (red lines), our model recovers the overall distribution of sending and receiving activities across the nodes. For example, node 1 and 10 show significantly higher level of both sending and receiving activities relative to the rest, and the model-simulated data captures those big jumps, without using node-specific indicators. Outdegree distribution of some low-activity nodes are not precisely recovered, however, indegree distribution looks much better. Since we use more information in the recipient selection process (i.e., network effects) while we only rely on minimum time-increments when choosing the observed sender, these results are expected. Lower left plot is the distribution of recipient sizes, where the x-axis spans over the size of recipients 1 to 14 (which is the maximum size of observed recipients) and the y-axis denotes the number of documents with x-number of recipients. The result shows that our model is underestimating single-receiver documents while overestimating documents with two, three, and four recipients. One explanation behind what we observe is that the model is trying to recover so-called ``broadcast'' emails, which are the documents with $\geq 10$ number of recipients, so that the intercept estimate $b_1$ is slightly moved toward right. To improve the goodness of fit for recipient sizes, we can add more covariates (e.g., sender indicators) or assign strong prior structure on $b_1$. In the end, the plot on the lower right is the PP plot for time-increments, which is a graphical measure commonly used for assessing how closely two data sets agree, by plotting the two cumulative distribution functions against each other. The resulting goodness of fit of the diagonal line gives a measure of the difference between the distribution of simulated time-increments and the observed ones, and our PP plot shows that we have great performance in reproducing the observed time distribution. All our findings from predictive experiments in Section \ref{subsec:Experiment_email} are further revealed in the PPC results from exponential distribution, and the plots comparing posterior checks between lognormal and exponential distributions are attached in Appendix B.

	\subsection{Parameter Estimates and Interpretation}\label{subsec:Result_email}
	Based on some preliminary experiments, we choose lognormal distribution to model time-to-next-email in Montgomery county email data, thus we estimate the variance parameter $\sigma_\tau^2$ with the prior specified as inverse-Gamma($2, 1$). In this section, we present the results based on 55,000 MCMC outer iterations with a burn-in of 15,000, where we thin by keeping every 40th sample. While the inner iterations for $\boldsymbol{\eta}$ and $\sigma_\tau^2$ are fixed as 1, we use 10 inner iterations for $\boldsymbol{b}$ to adjust for its slower convergence rate.
	\begin{figure}[!b]
		\centering
		\includegraphics[width=1\textwidth]{img/best-1.png}	
		\caption {Posterior distribution of $\boldsymbol{b}$ estimates.}
		\label{figure:betaresults}
	\end{figure}

@@

	To use this distribution efficiently, we derive a closed-form expression for $Z(\boldsymbol{\lambda}_{id})$ that does not require brute-force summation over the support of $\boldsymbol{u}_{ad}$ (\textit{i.e.} $\forall \boldsymbol{u}_{ad} \in [0,1]^A$). We recognize that if $\boldsymbol{u}_{ad}$ were drawn via independent Bernoulli distributions in which $P({u}_{adr}=1|\delta, \boldsymbol{\lambda}_{ad})$ was given by logit$(\lambda_{adr})$, then 
	\begin{equation*}
	P(\boldsymbol{u}_{ad}|\boldsymbol{\lambda}_{ad}) \propto \exp\Big\{\sum_{r \neq a } \lambda_{adr}u_{adr}\Big\}.  	 
	\end{equation*}
	This is straightforward to verify by looking at 
	\begin{equation*}
	&P(u_{adr}=1|\boldsymbol{u}_{ad[-r]}, \boldsymbol{\lambda}_{ad})
	=\frac{ \exp{(\lambda_{adr})}}{\exp{(\lambda_{adr})} + 1}.\end{aligned}\end{equation*}
	We denote the logistic-Bernoulli normalizing constant as $Z^{l}(\boldsymbol{\lambda}_{ad})$, which is defined as 
	\begin{equation*}
	Z^{l}(\delta,\boldsymbol{\lambda}_{ad})=\sum_{\boldsymbol{u}_{ad} \in [0,1]^{A}} \exp\Big\{\sum_{r\neq a} \lambda_{adr}u_{adr}\Big\}.
	\end{equation*}
	Now, since 
	\begin{equation*}
	&\exp\Big\{ \mbox{log}\Big(\text{I}(\lVert \boldsymbol{u}_{ad} \rVert_1 > 0)\Big) + \sum_{r \neq a} \lambda_{adr}u_{adr} \Big\}\\&= \exp\Big\{  \sum_{r \neq a} \lambda_{adr}u_{adr} \Big\},
	\end{equation*}
	except when $\lVert \boldsymbol{u}_{ad} \rVert_1=0$, we note that 
	\begin{equation*}
	Z(\boldsymbol{\lambda}_{ad})& = Z^{l}(\boldsymbol{\lambda}_{ad}) -\exp\Big\{ \sum\limits_{\forall u_{adr}=0}\lambda_{adr}u_{adr} \Big\}
	\\& = Z^{l}(\lambda_{a}^{(d)}) -  1.
	\end{equation*}
	We can therefore derive a closed form expression for $Z(\delta,\boldsymbol{\lambda}_{ad})$ via a closed form expression for $Z^{l}(\delta,\boldsymbol{\lambda}_{ad})$. This can be done by looking at the probability of the zero vector under the logistic-Bernoulli model:
	\begin{equation*}
	&\frac{\exp\Big\{ \sum\limits_{\forall u_{adr}=0}\lambda_{adr}u_{adr} \Big\}}{Z^{l}(\boldsymbol{\lambda}_{ad})}= \prod_{r \neq a}   \Big(1-\frac{ \exp{(\lambda_{adr})}}{\exp{(\lambda_{adr})} + 1}\Big).
	\end{equation*}
	Then, we have 
	\begin{equation*}
	& \frac{1}{Z^{l}(\boldsymbol{\lambda}_{ad})} &= \prod\limits_{r \neq a}\frac{1}{ \exp(\lambda_{adr})+ 1}.
	\end{equation*}
	Finally, the closed form expression for the normalizing constant is  \begin{equation*}
	\begin{aligned}Z(\boldsymbol{\lambda}_{ad}) = \prod_{r \neq a } \big(\mbox{exp}(\lambda_{adr}) + 1\big)-1.
	\end{equation*}
	\subsection*{Appendix B: Comparison of PPC Results: lognormal vs. exponential}\label{subsec: PPCexp}
	\iffalse
	\begin{figure}[H]
		\includegraphics[width=0.495\textwidth]{img/outdegree_exp-1.png}	
		\includegraphics[width=0.495\textwidth]{img/indegree_exp-1.png}	
		\includegraphics[width=0.495\textwidth]{img/recipientsize_exp-1.png}	
		\includegraphics[width=0.31\textwidth]{img/timedist_exp-1.png}
		\caption {PPC results from exponential distribution: outdegree distribution (\textit{upper left}), indegree distribution (\textit{upper right}), recipient size distribution (\textit{lower left}), and time-increments probability-probability (PP) plot  (\textit{lower right}). Red lines in the first three plot depict the observed statistics, and the red line in the last plot is the diagonal line connecting $(0, 0)$ and $(1, 1)$.}
		\label{figure:PPCexp}
	\fi
		\begin{figure}[H]
			\includegraphics[width=0.495\textwidth]{img/outdegree_two-1.png}	
			\includegraphics[width=0.495\textwidth]{img/indegree_two-1.png}	
			\includegraphics[width=0.495\textwidth]{img/recipientsize_two-1.png}	
			\includegraphics[width=0.31\textwidth]{img/timepp_two-1.png}
